```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
pre {
  max-height: 500px;
  overflow-y: auto;
}
pre[class] {
  max-height: 500px;
}
```

```{r,warning=F,message=F,echo=FALSE}
library(sgsR)
library(terra)

#--- Load mraster and access files ---#
r <- system.file("extdata", "wall_metrics_small.tif", package = "sgsR")

#--- load the mraster using the terra package ---#
mraster <- terra::rast(r)

a <- system.file("extdata", "roads.shp", package = "sgsR")

#--- load the access vector using the sf package ---#
access <- sf::st_read(a)

#--- apply kmeans algorithm to metrics raster ---#
sraster <- strat_kmeans(mraster = mraster, 
                        nStrata = 4) # algorithm will plot output

#--- apply kmeans algorithm to metrics raster ---#
existing <- sample_srs(raster = mraster, # use sraster as input for sampling
                       nSamp = 200, # request 200 samples be taken
                       mindist = 100) # algorithm will plot output

```

# `sample_*` functions {#sample .unnumbered}

A total of 5 `sample` functions currently exist in the package:

*   `sample_srs` - simple random sampling

*   `sample_grid`- grid sampling

*   `sample_strat` - stratified sampling within a `sraster`

*   `sample_clhs` - Latin hypercube sampling algorithm

*   `sample_balanced` - Balanced sampling [@R-BalancedSampling]

## Access {#access .unnumbered}

A feature of the `sample_*` functions that may be of interest is the ability to define `access`. Users can supply a road access network (must be `sf` line objects) and define buffers around `access` where samples should be excluded and included. 

Important additional parameters when `access` is provided are:

* `buff_inner` - An inner buffer that defines the distance from `access` where samples cannot be takes (i.e. if you do not want samples within 50 m of your `access` layer set `buff_inner = 50`). 

* `buff_outer` - Maximum distance samples can be located from `access` (i.e. if you dont want samples more than 200 meters from a road set `buff_inner = 200`)

## `sample_srs` {#srs .unnumbered}
We provided a very simple example of using the `sample_srs` function when we were preparing our [`existing`](#srast) data. We provide some additional examples below. 

::: {.box-content .vig-content}
::: {.box-title .vig-content-top}
### Input raster {.unnumbered}
:::
Notice that the input for `sample_srs` is `raster`. This means that either an `sraster` or an `mraster` can be supplied.
:::
 
```{r,warning=F,message=F} 
#--- perform simple random sampling ---#
sample_srs(raster = sraster, # input sraster
           nSamp = 200, # number of desired samples
           plot = TRUE) # plot
```

```{r,warning=F,message=F} 
sample_srs(raster = mraster, # input mraster
           nSamp = 200, # number of desired samples
           access = access, # define access road network
           mindist = 200, # minimum distance samples must be apart from one another
           buff_inner = 50, # inner buffer - no samples within this distance from road
           buff_outer = 200, # outer buffer - no samples further than this distance from road
           plot = TRUE) # plot
```

```{r,warning=F,message=F} 
sample_srs(raster = sraster, # input
           nSamp = 200, # number of desired samples
           access = access, # define access road network
           buff_inner = 50, # inner buffer - no samples within this distance from road
           buff_outer = 200, # outer buffer - no samples further than this distance from road
           plot = TRUE, # plot
           filename = tempfile(fileext = ".shp")) # write output samples to file
```

## `sample_grid` {#grid .unnumbered}

The `sample_grid` function applies regular grid sampling across an area where the `gridsize` parameter defines distance between samples.

```{r,warning=F,message=F} 
#--- perform grid sampling ---#
sample_grid(raster = sraster, # input sraster
            gridsize = 200, # grid distance
            plot = TRUE) # plot
```

```{r,warning=F,message=F} 
sample_grid(raster = sraster, # input sraster
            gridsize = 100, # grid distance
            access = access, # define access road network
            buff_inner = 50, # inner buffer - no samples within this distance from road
            buff_outer = 200) # outer buffer - no samples further than this distance from road
```

```{r,warning=F,message=F} 
sample_grid(raster = mraster, # input mraster
            gridsize = 200, # grid distance
            access = access, # define access road network
            buff_inner = 100, # inner buffer - no samples within this distance from road
            buff_outer = 400, # outer buffer - no samples further than this distance from road
            filename = tempfile(fileext = ".shp"), # write output samples to file
            plot = TRUE) # plot
```

## `sample_strat` {#sstrat .unnumbered}

The `sample_strat` function contains a hierarchical sampling algorithm originally developed my Martin Queinnec. The algorithm uses a moving window (`wrow` and `wcol` parameters) to filter the input `sraster` for locations where stratum pixels are spatially grouped rather than dispersed individually across the landscape.

The sampling is performed in 2 stages:

* **Rule 1** - Sample within spatially grouped stratum pixels.

* **Rule 2** - If no more samples exist to satisfy desired sampling count, individual stratum pixels are sampled.

The rule applied to a select a particular sample is defined in the `rule` attribute of output samples. We give a few examples below:

```{r,warning=F,message=F} 
#--- perform stratified sampling random sampling ---#
sample_strat(sraster = sraster, # input sraster
             nSamp = 200, # desired sample number
             plot = TRUE) # plot
```

In some cases, a user may wish to include an `existing` sample dataset within the algorithm. In order to adjust the total number of samples needed per stratum to reflect those already present in `existing`, we can use the utility function `extract_strata`.

This function takes an input `sraster` and the `existing` sample dataset and extracts the stratum for each sample. These samples can then be input into the `sample_strat` function which adjusts total required sample per class based on representation in `existing`.

```{r,warning=F,message=F} 
#--- extract strata values to existing samples ---#              
e.sr <- extract_strata(sraster = sraster, # input sraster
                       existing = existing) # existing samples to add strata value to

e.sr
```
::: {.box-content .vig-content}
::: {.box-title .vig-content-top}
### Existing strata {.unnumbered}
:::
Notice that `e.sr` now has an attribute named strata. If that parameter is not there, `sample_strat` will give an error.
:::

```{r,warning=F,message=F} 
sample_strat(sraster = sraster, # input sraster
             nSamp = 200, # desired sample number
             access = access, # define access road network
             existing = e.sr, # existing samples with strata values
             mindist = 200, # minimum distance samples must be apart from one another
             buff_inner = 50, # inner buffer - no samples within this distance from road
             buff_outer = 200, # outer buffer - no samples further than this distance from road
             plot = TRUE) # plot
```

::: {.box-content .vig-content}
::: {.box-title .vig-content-top}
### `mindist` {.unnumbered}
:::
The `mindist` parameter defined in the above example specifies the minimum euclidian distance that samples must be apat from one another.
:::

Notice that the sample outputs have `type` and [`rule`](#sstrat) attributes which outline whether samples are `existing` or `new` and whether `rule1` or `rule2` were used to select individual samples.

```{r,warning=F,message=F} 
sample_strat(sraster = sraster, # input
             nSamp = 200, # desired sample number
             access = access, # define access road network
             existing = e.sr, # existing samples with strata values
             include = TRUE, # include existing plots in nSamp total
             buff_inner = 50, # inner buffer - no samples within this distance from road
             buff_outer = 200, # outer buffer - no samples further than this distance from road
             filename = tempfile(fileext = ".shp"), # write output samples to file
             plot = TRUE) # plot
```

::: {.box-content .vig-content}
::: {.box-title .vig-content-top}
### `include` parameter {.unnumbered}
:::
The `include` parameter determines whether existing samples should be included in the total count of samples defined by `nSamp`. By defaults `include = FALSE`.
:::

## `sample_clhs` {#clhs .unnumbered}

The `sample_clhs` function implements conditioned latin hypercube (clhs) sampling functionality.  A number of other functions in the `sgsR` package help to provide guidance on clhs sampling including [`calculate_lhsPop`](#lhspop) and [`calculate_lhsOpt`](#lhsopt). Be sure to check out those functions for better understanding of how to optimize sample numbers.

Syntax for this function is similar to others shown above, though parameters like `iter`, which define the number of iterations within the Metropolis-Hastings process are important to consider. In these examples we use a low `iter` value because it takes less time to run. Default values for `iter` within the `clhs` package are 10,000.

```{r,warning=F,message=F} 
sample_clhs(mraster = mraster, # input
            nSamp = 200, # desired sample number
            plot = TRUE, # plot 
            iter = 100) # number of iterations
```


```{r,warning=F,message=F} 
sample_clhs(mraster = mraster, # input
            nSamp = 400, # desired sample number
            existing = existing, # existing samples
            iter = 100, # number of iterations
            details = TRUE) # clhs details
```

```{r,warning=F,message=F} 
sample_clhs(mraster = mraster, # input
            nSamp = 300, # desired sample number
            iter = 100, # number of iterations
            existing = existing, # existing samples
            access = access, # define access road network
            buff_inner = 100, # inner buffer - no samples within this distance from road
            buff_outer = 300, # outer buffer - no samples further than this distance from road
            plot = TRUE) # plot
```


::: {.box-content .vig-content}
::: {.box-title .vig-content-top}
### `cost` parameter {.unnumbered}
:::
The `cost` parameter defines an `mraster` covariate to be used to constrain the clhs sampling. This could be any number of variables. An example could be the distance a pixel is from road access (see example below), terrain slope, the output from [`calculate_coobs`](#coobs), or many others.
:::

```{r,warning=F,message=F} 
#--- cost constrained examples ---#
#--- calculate distance to access layer for each pixel in mr ---#
mr.c <- calculate_distance(raster = mraster, # input
                           access = access,
                           plot = TRUE) # define access road network

```

```{r,warning=F,message=F} 
sample_clhs(mraster = mr.c, # input
            nSamp = 250, # desired sample number
            iter = 100, # number of iterations
            cost = "dist2access", # cost parameter - name defined in calculate_distance()
            plot = TRUE) # plot
```

```{r,warning=F,message=F} 
sample_clhs(mraster = mr.c, # input
            nSamp = 250, # desired sample number
            existing = existing, # existing samples
            iter = 100, # number of iterations
            cost = "dist2access", # cost parameter - name defined in calculate_distance()
            plot = TRUE) # plot
```

## `sample_balanced` {#balanced .unnumbered}

The `sample_balanced` algorithm performs the balanced sampling methodology from the `stratifyR` and `SamplingBigData` packages. These functions can take a bit longer to run so be patient!

```{r,warning=F,message=F} 
sample_balanced(mraster = mraster, # input
                nSamp = 200, # desired sample number
                plot = TRUE) # plot
```

```{r,warning=F,message=F} 
sample_balanced(mraster = mraster, # input
                nSamp = 100, # desired sample number
                algorithm = "lcube", # algorithm type
                access = access, # define access road network
                buff_inner = 50, # inner buffer - no samples within this distance from road
                buff_outer = 200) # outer buffer - no samples further than this distance from road
```