```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
pre {
  max-height: 500px;
  overflow-y: auto;
}
pre[class] {
  max-height: 500px;
}
```

```{r,warning=F,message=F,echo=FALSE}
library(sgsR)
library(terra)

#--- Load mraster and access files ---#
r <- system.file("extdata", "wall_metrics_small.tif", package = "sgsR")

#--- load the mraster using the terra package ---#
mraster <- terra::rast(r)

a <- system.file("extdata", "roads.shp", package = "sgsR")

#--- load the access vector using the sf package ---#
access <- sf::st_read(a)

#--- apply kmeans algorithm to metrics raster ---#
sraster <- strat_kmeans(mraster = mraster, 
                        nStrata = 4) # algorithm will plot output

#--- apply kmeans algorithm to metrics raster ---#
existing <- sample_srs(raster = mraster, # use sraster as input for sampling
                       nSamp = 200, # request 200 samples be taken
                       mindist = 100) # algorithm will plot output

```


# `calculate_*` functions {#calc .unnumbered}

A total of 7 `calculate` functions currently exist in the package:

-   `calculate_distance` - per pixel distance to closes access vector

-   `calculate_pcomp`- calculate principal components on input `mraster`

-   `calculate_reqSamples` - calculate the number of samples needed for represntative sample within a `sraster`

-   `calculate_ahels` - appy the adapted Hypercube Evaluation of a Legacy Sample (ahels) algorithm to existing samples

-   `calculate_coobs` - apply count of observations algorithm to existing samples

-   `calculate_lhsPop` - calculate population covariate statistics for latin hypercube sampling

-   `calculate_lhsOpt` - calculate optimal latin hypercube sampling paramters including sample number
