--- 
title: "Structurally Guided Sampling"
author: "Tristan Goodbody"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [sgs.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview {-#overview}

This document is a current summary and explaination of the structurally guided sampling ([`sgsR`](https://github.com/tgoodbody/sgsR)) project being undertaken by [Dr. Tristan R.H. Goodbody](https://scholar.google.ca/citations?user=C_qdYDsAAAAJ&hl=en) and [Dr. Nicholas C. Coops](https://scholar.google.ca/citations?user=XvUF8dUAAAAJ&hl=en) at the IRSS Lab at the UBC Faculty of Forestry. 

Herein we provide a:

* Literature review about [SGS](#SGS)

* [Describe package fundamentals](#fund)

* Overview of [stratification algorithms](#strat)

* Overview of [sampling algorithms](#sample)

* Overview of [calculate algorithms](#calc)

To install the package from the github please run the following.

```{r message = FALSE, warning = FALSE, eval = FALSE}
install.packages("devtools")
devtools::install_github("https://github.com/tgoodbody/sgsR", build_vignettes = FALSE)
library(sgsR)
```
 
